---
title: "p8105_hw1_ns3923"
author: Nichcha Subdee (ns3923)
output: github_document
---

## Problem 1
This problem uses inline R, plotting, and shows how `ggplot` behaves with different variable types. I first load the necessary packages.

```{r, message=FALSE}
library(moderndive)
library(ggplot2)
```

Then, we will load the `early_january_weather` dataset, which contains hourly meteorological data for LGA, JFK, and EWR for January 2013.

```{r}
data("early_january_weather")
```

Next, we would like to know the variables and the size of the dataset and the mean temperature. We can use the following code to do so.

```{r, eval=FALSE}
# The variables in this dataset: Names and values
names(early_january_weather)
?early_january_weather

# The size of the dataset as numbers of rows and columns, respectively
nrow(early_january_weather)
ncol(early_january_weather)

# The mean temperature in 2 decimal points
round(mean(early_january_weather$temp, na.rm = TRUE),2)
```
**Discussion**

The dataset has `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns.  

There are **15 variables**: `r names(early_january_weather)`

The variable `origin` indicates the airport weather station. Temperature (`temp`) and dewpoint (`dewp`) are recorded in degrees Fahrenheit, with a mean temperature of **`r round(mean(early_january_weather$temp, na.rm = TRUE),2)` °F.** Other important variables include relative humidity (`humid`), wind direction (`wind_dir`, in degrees), wind speed and gust (`wind_speed`,`wind_gust`, in mph), precipitation (`precip`, in inches), sea level pressure (`pressure`, in millibars), and date and hour of the recording (`time_hour`).

Next, we will create a scatterplot of temp (y) vs time_hour (x) by using code as follow.

```{r}
# Create scatterplot
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point()

```

**Scatterplot explanation**:

The scatterplot shows that temperature fluctuates over time in early January 2013. The color scale shows humidity, which is generally higher when temperatures are lower.

We can export the plot by using the following code
```{r, eval=FALSE}
# Save plot
ggsave("scatterplot_temp_time_hour.png")
```

## Problem 2
Create a dataframe comprised of 

*  a random sample of size 10 from a standard Normal distribution
*  a logical vector indicating whether elements of the sample are greater than 0
*  a character vector of length 10
*  a factor vector of length 10, with 3 different factor “levels”

Creating the dataframe

```{r setup, include=FALSE}
# Load tidyverse library
library(tidyverse)
```

```{r df_create}
# Fix the random sample
set.seed(1234)

df = tibble(
  num = rnorm(10),
  logic = num > 0,
  char = letters[1:10],
  fact = factor(rep(c("low","medium","high"), length.out = 10))
)

df
```

Try to take **the mean of each variable** in your dataframe. 

```{r variable mean}
mean(pull(df, num))
mean(pull(df, logic))
mean(pull(df, char))
mean(pull(df, fact))
```

**What works and what doesn’t?**

The mean of num and logic can be computed without issue. The numeric variable `num` behaves as expected. The logical variable `logic` also works because R interprets TRUE as 1 and FALSE as 0. In contrast, the means of `char` and `fact` cannot be calculated. Character variables are non-numeric text values, and factor variables represent categorical labels. This causes their means to be undefined in this context, since they cannot be directly converted into meaningful numeric values.

















